---
title: "Musclestemcells"
author: "Srikanth Ravichandran"
date: "24 March 2020"
output: html_document
---

# Computational analyses based on 'differential expression' to predict intracellular signaling molecules involved in different subpopulations of CD34+ cells 

```{r include=FALSE,echo=T}
# Required libraries
  library("topGO") 
  library (plyr)
  #library(SigHotSpotter)
  library(plyr)
  library(igraph)
  library(Matrix)
  library(reshape2)
  library(parallel)
  library(RSpectra)
  library(dplyr)
  library(snow)
  library(backports)
  library(DESeq2)
  library(zinbwave)
  library(data.table)
  library(visNetwork)
  source("cd34proj_allfunctions.R")
  source("main_library.R")
  source("for_plotting_network_functions.R")
```

## Load data and preprocessing
```{r,message=FALSE}
load("MOUSE_Background_Network_omnipath_reactome_metacore_01042019.RDATA")
# Reading the excel sheest with data and annotation
# The command uses a function to read all sheets as a lists
data<-read_excel_allsheets("Heterogeneity_CD34_FOXO_project_Data.xlsx",F)
tflist<-read.table("Mus_musculus_TF.txt",sep="\t",header=T)
# extracting the data from excel WB
annotation <- data$`Sample names`
expressionData <- join_all(c(data[3:5]),by=c("Gene"))
#saving the data as r.data
save(annotation,expressionData, file = "CD34ExpAndAnnodata.RData")
# load the CD34 data
# load("C:/Users/olive/OneDrive/DesktopMY/CELIUM_BIO/mycelial_scaffold/models/signalingfactorscd34positive-master/CD34ExpAndAnnodata.RData")
# read differentiall expression data
```
### Extraction of different cell subpopulations
```{r}
# to grep data
Young_CD34High<-cbind(expressionData[1],expressionData[,c(grep("Young_CD34High",annotation$Name,value=T))]) #combining from set 7,8,9 and 10
Young_CD34High_set7_8<-Young_CD34High[,1:8] # data from only set 7 and 8
Young_CD34High_set7_8_9<-Young_CD34High[,1:12] # data from only set 7 and 8 and 9
Young_CD34High_set9<-Young_CD34High[,c(1,9:12)] # data from only set 7 and 8
Young_CD34Low<-cbind(expressionData[1],expressionData[,c(grep("Young_CD34Low",annotation$Name,value=T))]) #combining from set 7,8,9 and 10
Old_CD34High<-cbind(expressionData[1],expressionData[,c(grep("Old_CD34High",annotation$Name,value=T))])
Old_CD34Low<-cbind(expressionData[1],expressionData[,c(grep("Old_CD34Low",annotation$Name,value=T))])
Geriatric_CD34High<-cbind(expressionData[1],expressionData[,c(grep("Geriatric_CD34High",annotation$Name,value=T))])
Geriatric_CD34Low<-cbind(expressionData[1],expressionData[,c(grep("Geriatric_CD34Low",annotation$Name,value=T))])
```
### Read and process differential expression data
Extract differentially expressed genes. Here, all significantly differentially expressed TFs are taken without any log-fold change cutoff.
```{r,message=FALSE}
deExpData<-read_excel_allsheets("Signatures_CD34.xlsx",F)
# Extract differentially expressed genes. Here, all significantly differentially expressed TFs are taken without any log-fold change cutoff.
boolDeExp=lapply(deExpData,cd34deBool)
# shortlisting highly significant DEtfs as there are many TFs differentially expressed for young vs geraitic
YhivsGhi0.000002<-deExpData$`High DE Young v Geriatric`[deExpData$`High DE Young v Geriatric`$padj<2.07242382665159E-07,] # genes shortlisted by the bioinformatician as signatures
YhivsGhi0.000002Bool<-cd34deBool(YhivsGhi0.000002)
names(boolDeExp)
```

### 1. "Young Signatures DE HvL"
```{r}

res1_set7_8<-runSigHotSpotter(Young_CD34High_set7_8,Young_CD34Low,100,90,acc=boolDeExp$`Young Signatures DE HvL`)

```

### 6. "High DE Young v Geriatric"
```{r}

res6_set7_8_9<-runSigHotSpotter(Young_CD34High_set7_8_9,Geriatric_CD34High,100,90,acc=YhivsGhi0.000002Bool)


```

# Results of Analysis III
# Identification of signaling hotspots Considering Foxos as targets
### Reading refbool discritized data
```{r}
foxTfs <- data.frame(Gene    = c("Foxo1","Foxo3","Foxo4"),
                     BoolExp = rep(1, 3),
                     stringsAsFactors = FALSE)
# foxTfs<-read.table("C:/Users/olive/OneDrive/Desktop/MYCELIUM_BIO/mycelial_scaffold/models/signalingfactorscd34positive-master/data/foxtfs.txt",sep="\t",header = T)
colnames(foxTfs)<-c("Gene","BoolExp")
#adding foxogenes as non_interface_TFs
non_interface_TFs<-rbind(non_interface_TFs,foxTfs[1])
```
# Finding signaling hotspots considering Foxos (Foxo1,Foxo3 and Foxo4 are considered to be upregulated) as terminal TFs
```{r}
#Change the SigHotSpotter_standalone to include Foxos as non-interface TFs. Change non_interface_TFs -> non_interface_TFsF
res8foxo<-runSigHotSpotterFoxo(Young_CD34High_set7_8,Young_CD34Low,100,90,acc=foxTfs)
```
# for plotting the dotplots of final results
```{r}
#' function to plot
#' @param x is the res1[[1]]
plotBarResult<-function(x){
a=x$trimmed_score_A
b=x$trimmed_score_I
names(a)[1]=c("Gene")
names(b)[1]=c("Gene")
d=rbind(a,b)
d=d[order(d$`Compatibility score`),]
d$color[d$`Compatibility score`>0.5]="darkgreen"
d$color[d$`Compatibility score`<0.5]<-"red"
d$group[d$`Compatibility score`>0.5]='Active'
d$group[d$`Compatibility score`<0.5]<-'Inactive'
dotchart(d$`Compatibility score`,labels = d$Gene,cex=0.8,groups = factor(d$group),xlim=c(min(d$`Compatibility score`)-0.005,max(d$`Compatibility score`)+0.04),main="Signaling Hotspots",xlab="Compatibility score",color=d$color,gcolor="black")
# Adding text
text(x = d$`Compatibility score`, y = 1:nrow(d)+ifelse(d$group=="Active",2,0), labels= d$`Compatibility score`,cex = 0.6,pos = 4)
}
```

# Plot results
```{r}
plotBarResult(res1_set7_8[[1]])
plotBarResult(res1_set7_8[[2]])

plotBarResult(res6_set7_8_9[[1]])
plotBarResult(res6_set7_8_9[[2]])

plotBarResult(res8foxo[[1]])
plotBarResult(res8foxo[[2]])
```

#### Saving network files for plotting in Cytoscape
```{r eval=FALSE, include=FALSE}
youngcd34hivsLownodes<-res1_set7_8[[1]]$vis_net_I[[which(res1_set7_8[[1]]$trimmed_score_I=="Akt1")]]$nodes
youngcd34hivsLowedges<-res1_set7_8[[1]]$vis_net_I[[which(res1_set7_8[[1]]$trimmed_score_I=="Akt1")]]$edges
youngcd34hivsGeraitricnodes<-res6_set7_8_9[[1]]$vis_net_I[[which(res6_set7_8_9[[1]]$trimmed_score_I=="Akt1")]]$nodes
youngcd34hivsGeraitricedges<-res6_set7_8_9[[1]]$vis_net_I[[which(res6_set7_8_9[[1]]$trimmed_score_I=="Akt1")]]$edges
factorsForFoxoNodes<-res8foxo[[2]]$vis_net_A[[which(res8foxo[[2]]$trimmed_score_A=="Akt1")]]$nodes
factorsForFoxoEdges<-res8foxo[[2]]$vis_net_A[[which(res8foxo[[2]]$trimmed_score_A=="Akt1")]]$edges

# writing the output network tables
write.table(youngcd34hivsLownodes,"youngcd34hivsLownodes.txt",sep="\t",row.names = F,quote = F)
write.table(youngcd34hivsLowedges,"youngcd34hivsLowedges.txt",sep="\t",row.names = F,quote = F)
write.table(youngcd34hivsGeraitricnodes,"youngcd34hivsGeraitricnodes.txt",sep="\t",row.names = F,quote = F)
write.table(youngcd34hivsGeraitricedges,"youngcd34hivsGeraitricedges.txt",sep="\t",row.names = F,quote = F)
write.table(factorsForFoxoNodes,"factorsForFoxoNodes.txt",sep="\t",row.names = F,quote = F)
write.table(factorsForFoxoEdges,"factorsForFoxoEdges.txt",sep="\t",row.names = F,quote = F)
```
